---
title: "TIM2024Analysis"
author: "Zach Peagler"
date: "2024-05-03"
output: html_document
---

# Introduction

This trial was conducted in a greenhouse at Kennesaw State University (KSU). Plants were assigned treatments in an offset grid pattern. Height measurements were taken regularly, and fluorescence measurements were conducted biweekly starting when the plant had enough leaf material to analyze with an Li-600.

# 1.Setup

## 1.1 Load Packages
```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse); require(lmerTest); require(showtext); require(MASS); require(scico); require(pwr); require(devtools); require(multcomp); require(rstatix); require(car); require(vegan)
# custom package
install_github("zachpeagler/ztils")
require(ztils)
```

## 1.2 Initialize Variables
```{r graphical_setup}

a_palette <- "oslo"
# super lazy way to make palettes
two_colors = scico(3, palette=a_palette)
four_colors = scico(5, palette=a_palette)
five_colors = scico(6, palette=a_palette)
true_two_col = scico(2, palette=a_palette)
ten_col = scico(10, palette=a_palette)
twelve_col = scico(13, palette=a_palette)

# add fonts
font_add_google("Open Sans", family = "open")
font_add_google("Montserrat", family = "mont")
# initialize imported fonts
showtext_auto()
# font sizes
## title
title_size = 16
## subtitle
subtitle_size = 12
## caption
caption_size = 10
## text
text_size = 10

# custom shapes
four_shapes = c(15,16,17,23)

##### Experimental Constants

tim_germ_date <- "2024-04-01"

tim_treatment_order <- c("Control",
                     "Liquid",
                     "CG",
                     "CBG")

tim_levels <- c("A1", "A2", "A3", "A4", "A5", "A6",
                   "A7", "A8", "A9", "A10", "A11", "A12", "A13",
                   "B1", "B2", "B3", "B4", "B5", "B6",
                   "B7", "B8", "B9", "B10", "B11", "B12", "B13",
                   "C1", "C2", "C3", "C4", "C5", "C6",
                   "C7", "C8", "C9", "C10", "C11", "C12", "C13",
                   "D1", "D2", "D3", "D4", "D5", "D6",
                   "D7", "D8", "D9", "D10", "D11", "D12", "D13",
                   "E1", "E2", "E3", "E4", "E5", "E6",
                   "E7", "E8", "E9", "E10", "E11", "E12",
                   "F1", "F2", "F3", "F4", "F5", "F6",
                   "F7", "F8", "F9", "F10", "F11", "F12",
                   "G1", "G2", "G3", "G4", "G5", "G6",
                   "G7", "G8", "G9", "G10", "G11", "G12",
                   "H1", "H2", "H3", "H4", "H5", "H6",
                   "H7", "H8", "H9", "H10", "H11", "H12"
                )

```

## 1.3 Load and Clean Data

### 1.3.1. Fluoro Data
``` {r fluoro_data, include = TRUE}

li_data_file <- "C:/Github/Thesis/data/TIM/TIM24_Fluoro.csv"
li_data <- read.csv(li_data_file)[,c(2,3,7,8,9,10,17,28,32,34,35,36,39,40)] %>%
  filter(leak_pct<10 & gsw > 0) %>%
  mutate(Date = parse_date_time(Date, orders = "mdy"),
         DaysFromGermination = as.numeric(round(difftime(Date, tim_germ_date, units = c("days")), 0)),
         Date = as.factor(as.Date(Date)),
         Time = parse_date_time(Time, orders = "T"),
         Time = as.factor(format(Time, "%H:%M:%S")),
         Treatment = factor(Treatment, 
                            levels = tim_treatment_order),
         Inoculation = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~TRUE,
           Treatment=="CG"~FALSE,
           Treatment=="CBG"~TRUE
         ),
         Chitosan = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~FALSE,
           Treatment=="CG"~TRUE,
           Treatment=="CBG"~TRUE
         ),
         Column = case_when(
           Column==1~"A",
           Column==2~"B",
           Column==3~"C",
           Column==4~"D",
           Column==5~"E",
           Column==6~"F",
           Column==7~"G",
           Column==8~"H",
         ),
         Plant = paste0(Column, Row),
         LogitPhiPS2 = logit(PhiPS2, FALSE),
         Column = as.factor(Column),
         Row = as.factor(Row),
         Plant = as.factor(Plant),
         Plant = factor(Plant, levels = tim_levels),
         P_atm = P_atm * 100
  ) %>%
  rename(Pot=Column)

data_tim_fluoro <- li_data[,c(5,16,17,1,2,15,3,4,18,9,10,12,14,11,6,8,19)] %>%
  rename(AmbientHumidity = rh_s,
         AmbientTemperature = Tref,
         AmbientPressure = P_atm,
         AmbientLight = Qamb,
         LeafTemperature = Tleaf
         )


```

### 1.3.2. Height Data
``` {r height_data, include = TRUE}

h_data_file <- "C:/Github/Thesis/data/TIM/TIM24_Height.csv"
h_data <- read.csv(h_data_file) %>%
  mutate(Date = parse_date_time(Date, orders = "mdy"),
         DaysFromGermination = as.numeric(round(difftime(Date, tim_germ_date, units = c("days")), 0)),
         Date = as.factor(as.Date(Date)),
         Column = case_when(
           Column==1~"A",
           Column==2~"B",
           Column==3~"C",
           Column==4~"D",
           Column==5~"E",
           Column==6~"F",
           Column==7~"G",
           Column==8~"H",
         ),
         Inoculation = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~TRUE,
           Treatment=="CG"~FALSE,
           Treatment=="CBG"~TRUE
         ),
         Chitosan = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~FALSE,
           Treatment=="CG"~TRUE,
           Treatment=="CBG"~TRUE
         ),
         Treatment = factor(Treatment, levels = tim_treatment_order),
         Plant = as.factor(paste0(Column, Row)),
         Plant = factor(Plant, levels = tim_levels),
         Row = as.factor(Row),
         Column = as.factor(Column)
  ) %>%
  rename(Pot = Column)

data_tim_height <- h_data[,c(3,7,8,4,6,1,2,9,5)]

```

### 1.3.3. Destructive Sampling Data
```{r DS_data, include=TRUE}
DS_data_file <- "C:/Github/Thesis/data/TIM/TIM24_DS.csv"
DS_data <- read.csv(DS_data_file) %>%
  mutate(
        Column = case_when(
           Column==1~"A",
           Column==2~"B",
           Column==3~"C",
           Column==4~"D",
           Column==5~"E",
           Column==6~"F",
           Column==7~"G",
           Column==8~"H",
         ),
         Inoculation = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~TRUE,
           Treatment=="CG"~FALSE,
           Treatment=="CBG"~TRUE
         ),
         Chitosan = case_when(
           Treatment=="Control"~FALSE,
           Treatment=="Liquid"~FALSE,
           Treatment=="CG"~TRUE,
           Treatment=="CBG"~TRUE
         ),
         Treatment = factor(Treatment, levels = tim_treatment_order),
         Plant = as.factor(paste0(Column, Row)),
         Plant = factor(Plant, levels = tim_levels),
         Row = as.factor(Row),
         Column = as.factor(Column)
  ) %>%
  rename(Pot = Column)

data_tim_ds <- DS_data[,c(3,10,11,1,2,12,4,5,6,7,8,9)]

```

# 2. Analysis

## 2.1. Fluorometric Parameters

### 2.1.1. Fluoro Exploratory Analysis
``` {r fluoro_exploratory_analysis, include=TRUE}
#gsw
multiPDF_Z(filter(li_data, gsw >0)$gsw, 100, "all", a_palette, "gsw")
multiCDF_Z(filter(li_data, gsw >0)$gsw, 100, "all", a_palette, "gsw")
multiKS_cont(filter(li_data, gsw >0)$gsw, "all")

###### Not significantly different than a Gamma or Normal distribution.
###### Try models with both.

# check for heteroscedasticity
## in gsw as a function of treatment
leveneTest(li_data$gsw~li_data$Treatment)
## in gsw as a function of time from start
leveneTest(li_data$gsw~li_data$MinutesFromStart)

###### Unequal variances. Uh oh.

## box plot by Treatment boxplot
ggplot(data = li_data, aes(x= Treatment, y = gsw, fill=Treatment, color=Treatment)) +
  geom_boxplot(alpha=.5, width=0.3, outliers = FALSE)+
    geom_jitter( width=.15, height=0)+
  scale_fill_manual(values=four_colors)+
  scale_color_manual(values=four_colors)+
  labs(
    title=str_wrap("Stomatal Conductance Across Inoculation Treatments in Tomato", 40)
  ) +
  ylab(str_wrap("Stomatal Conductance (mol m-2 s-1)", 20))+
#  annotate("text", x=1:4, y=3, label = c("a", "b", "b", "b"), size=5, family="open")+
  theme_bw()+
  theme(
    legend.position="none",
    text = element_text(size=16, family="mont", lineheight=0.8),
    axis.title = element_text(size=16, family = "mont", face= "bold"),
    title = element_text(size=20, family="open", face="bold")
  )


# GSW by date across relative humidity
ggplot(data = li_data, aes(x=Date, y = gsw, 
                           color = rh_s)) +
  geom_jitter(width=1, size=2, alpha=1)+
  scale_color_scico(begin=0.9, end=0, palette=a_palette)+
#  scale_x_discrete(guide=guide_axis(check.overlap=TRUE))+
  theme_bw()+
  ylab(str_wrap("Stomatal Conductance (mol m-2 s-1)", 20))+
  xlab("Days From Germination")+
  labs(
        tag="2023",
    title=str_wrap("Stomatal Conductance By Days From Germination Across Relative Humidity in Tomato", 50)
  )+
    guides(color=guide_colorbar(title=str_wrap("Relative Humidity", 10)))+
  theme(
    text = element_text(size=16, family="mont"),
    legend.title = element_text(size=16, family="mont", face="bold"),
    legend.text = element_text(size=14, family="mont"),
    legend.position="right",
    legend.title.position = "top",
    legend.key.height = unit(.4, "cm"),
    legend.background = element_rect(color=two_colors[2], fill=NA,
                                     linewidth=.5, linetype = 2),
    axis.title = element_text(size=16, family = "mont", face= "bold"),
    title = element_text(size=20, family="open", face="bold", lineheight = .8)
  )


# GSW by time across relative humidity
ggplot(data = li_data, aes(x=Time, y = gsw, 
                           color = rh_s,
                           shape = Treatment)) +
  geom_jitter(width=1, size=2, alpha=1)+
  scale_color_scico(begin=0.9, end=0, palette=a_palette)+
  scale_shape_manual(values=four_shapes)+
  scale_x_discrete(guide=guide_axis(check.overlap=TRUE))+
  theme_bw()+
  ylab(str_wrap("Stomatal Conductance (mol m-2 s-1)", 20))+
  xlab("Time")+
  labs(
    title=str_wrap("Stomatal Conductance By Time Across Date in Tomato", 30)
  )+
    guides(color=guide_legend(title=str_wrap("Relative Humidity", 10)))+
  theme(
    text = element_text(size=16, family="mont", lineheight = 1),
    legend.title = element_text(size=16, family="mont", face="bold"),
    legend.text = element_text(size=14, family="mont"),
    legend.position="right",
    legend.title.position = "top",
    legend.key.height = unit(.3, "cm"),
    legend.background = element_rect(color=two_colors[2], fill=NA,
                                     linewidth=.5, linetype = 2),
    axis.title = element_text(size=16, family = "mont", face= "bold"),
    title = element_text(size=20, family="open", face="bold", lineheight = .8)
  )

# gsw by Minutes from Start
ggplot(data = li_data, aes(x=MinutesFromStart, y = gsw, shape = Treatment, color = rh_s)) +
  geom_point(size=3)+
  scale_color_scico(begin=0.9, end=0, palette=a_palette)+
  theme_bw()+
  ylab(str_wrap("Stomatal Conductance (mol m-2 s-1)", 20))+
  xlab("Minutes From Start")+
  labs(
    title=str_wrap("Stomatal Conductance by Minutes From Start Across Relative Humidity and Treatment in Tomato", 40)
  )+
      guides(color=guide_colorbar(title=str_wrap("Relative Humidity", 8)))+
  theme(
    text = element_text(size=16, family="mont"),
    legend.position="right",
    axis.title = element_text(size=16, family = "mont", face= "bold"),
    title = element_text(size=20, family="open", face="bold", lineheight = .8)
  )

## phiPS2
### because phiPS2 is a proportion [0:1], we don't have to create PDFs and CDFs or perform KS tests. For a proportion we can just logit transform it then use a linear model.

### check for heteroscedasticity
#### Pr(>F) of < 0.05 means that the value is homoscedastic
leveneTest(li_data$PhiPS2~li_data$Treatment)

###### Heteroscedastic by treatment.

### box plot of logit PhiPS2 by treatment
ggplot(data = li_data, aes(x= Treatment, y = PhiPS2, 
                         fill=Treatment, color=Treatment)) +
  geom_boxplot(alpha=.5, width=0.5, outliers = FALSE)+
  geom_jitter( width=.15, height=0)+
  scale_fill_manual(values=four_colors)+
  scale_color_manual(values=four_colors)+
  labs(
    title=str_wrap("PhiPS2 by Inoculation Treatment in Tomato", 30)
  ) +
  ylab("PhiPS2")+
  xlab("Treatment")+
  theme_bw()+
  theme(
    legend.position="none",
    text = element_text(size=subtitle_size, family="mont", lineheight=0.8),
    axis.title = element_text(size=subtitle_size, family = "mont", face= "bold"),
    title = element_text(size=title_size, family="open", face="bold")
  )

### logit PhiPS2 by time scatterplot
ggplot(data = li_data, aes(x= MinutesFromStart, y = logitPS2, shape = Treatment, color=Qamb)) +
  geom_jitter( width=0, height=0)+
  scale_fill_manual(values=four_colors)+
  scale_color_scico(begin=0.9, end=0, palette=a_palette)+
#  scale_x_discrete(guide=guide_axis(check.overlap=TRUE))+
  labs(
    title=str_wrap("Logit PhiPS2 by Minutes From Start and Inoculation Treatment in Tomato", 40)) +
  ylab("PhiPS2")+
  xlab("MinutesFromStart")+
  theme_bw()+
  theme(
    legend.position="right",
    text = element_text(size=subtitle_size, family="mont", lineheight=0.8),
    axis.title = element_text(size=subtitle_size, family = "mont", face= "bold"),
    title = element_text(size=title_size, family="open", face="bold")
  )

```

### 2.1.2. Fluoro Models and Post Hoc Analysis
```{r fluoro_models}
## gsw could be normal or gamma
model_gsw<- lm(log(gsw) ~ Treatment + rh_s, data = li_data)
AIC(model_gsw)
summary(model_gsw)

predict_plot(model_gsw, li_data, gsw, rh_s, Treatment,
             correction = "exponential")

summary(model_fl_phi2<- lm(
  logitPS2 ~ Treatment + Date,
  data = li_data))

```

### Height Analysis

``` {r height_analysis, include = TRUE}

# height
multiPDF_Z(filter(h_data, Height >0)$Height, 100, "all", a_palette, "Height")
multiCDF_Z(filter(h_data, Height >0)$Height, 100, "all", a_palette, "Height")
multiKS_cont(filter(h_data, Height >0)$Height, "all")


### box plot of height by treatment
ggplot(data = h_data, aes(x= Treatment, y = Height, 
                         fill=Treatment, color=Treatment)) +
  geom_boxplot(alpha=.5, width=0.5, outliers = FALSE)+
  geom_jitter( width=.15, height=0)+
  scale_fill_manual(values=four_colors)+
  scale_color_manual(values=four_colors)+
  labs(
    title="Height by Inoculation Treatment in Tomato"
  ) +
#    annotate("text", x=1:4, y=35, label = c("a", "ab", "ab", "b"), size=6, family="open")+
  ylab("Height (cm)")+
  xlab("Treatment")+
  theme_bw()+
  theme(
    legend.position="none",
    text = element_text(size=subtitle_size, family="mont", lineheight=0.8),
    axis.title = element_text(size=subtitle_size, family = "mont", face= "bold"),
    title = element_text(size=title_size, family="open", face="bold")
  )

# height scatterplot
ggplot(data=h_data, aes(x=Date, y=Height, color=Treatment))+
  geom_jitter(size=2)+
  facet_wrap(~Treatment)+
  scale_color_manual(values = four_colors)+
  labs(
    title="Height by Date and Treatment in Tomato"
  )+
  ylab("Height (cm)")+
  theme_bw()+
    theme(
    legend.position="right",
    text = element_text(size=subtitle_size, family="mont", lineheight=0.8),
    axis.title = element_text(size=subtitle_size, family = "mont", face= "bold"),
    title = element_text(size=title_size, family="open", face="bold")
  )

summary(model_h <- lm(
  Height ~ Treatment,
  data = h_data))

comps <- glht(model_h, linfct = mcp(Treatment = "Tukey"))
cld(comps)

```

### Destructive Sampling Analysis

``` {r DS_analysis, include = TRUE}
# RS_length
multiPDF_Z(DS_data$RS_Length, 100, "all", a_palette, "R:S Length")
multiCDF_Z(DS_data$RS_Length, 100, "all", a_palette, "R:S Length")
multiKS_cont(DS_data$RS_Length, "all")


ggplot(data=DS_data, aes(x=Treatment, y=AG_Length, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
  labs(
    title="Aboveground Length by Treatment"
  )

ggplot(data=DS_data, aes(x=Treatment, y=AG_Mass, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
    labs(
    title="Aboveground Mass by Treatment"
  )

ggplot(data=DS_data, aes(x=Treatment, y=BG_Length, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
    labs(
    title="Belowground Length by Treatment"
  )

ggplot(data=DS_data, aes(x=Treatment, y=BG_Mass, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
      labs(
    title="Belowground Mass by Treatment"
  )

ggplot(data=DS_data, aes(x=Treatment, y=RS_Length, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
          labs(
    title="Root:Shoot Length by Treatment"
  )

ggplot(data=DS_data, aes(x=Treatment, y=RS_Mass, color=Treatment))+
  geom_violin()+
  geom_boxplot(width=0.2)+
  geom_jitter()+
        labs(
    title="Root:Shoot Mass by Treatment"
  )


summary(model_RSL <- (lm(
  logit(RS_Length) ~ Treatment,
  data = DS_data)))

summary(model_RSM <- (lm(
  logit(RS_Mass) ~ Treatment,
  data = DS_data)))

p1 <- rda(DS_data[,4:7])
summary(p1)
biplot(p1)


tr <- sort(unique(DS_data$Treatment))
Treatments <- tr[match(DS_data$Treatment, tr)]

cols <- c("red", "pink", "lightblue", "blue")

px <- scores(p1)$sites
vx <- scores(p1)$species

ggplot()+
  geom_point(data=px, aes(x=PC1, y=PC2, color=Treatments, shape=Treatments, fill=Treatments), size=3)+
  scale_color_manual(values = four_colors)+
  scale_fill_manual(values = four_colors)+
  scale_shape_manual(values = four_shapes)+
  geom_segment(data = vx, aes(x=0, y=0, xend=PC1*.225, yend=PC2*.225), color = "black")+
  annotate("text", x=vx[,1]*.25, y=vx[,2]*.25, label = rownames(vx))+
  labs(title=str_wrap("Principal Component Analysis for Tomato Destructive Sampling Metrics", 40))+
  xlab("PC1 (67.76%)")+
  ylab("PC2 (31.41%)")+
  theme_bw()+
  theme(
    text = element_text(size=subtitle_size, family="mont", lineheight=0.8),
    axis.title = element_text(size=subtitle_size, family = "mont", face= "bold"),
    title = element_text(size=title_size, family="open", face="bold")
  )

```